Adaptive Shrinkage and False Discovery Rates by Laplace Approximation

### 2013-11-07 paper/main.tex
add some R code for paper, and update some of the options

\def\x{\mbf{x}}
\def\y{\mbf{y}}
\def\e{\mbf{e}}
\def\g{\mbf{g}}
\def\p{p}
\def\Yb{\hat{Y}_1^{\text{BAYES}}}
\def\Xperp{X_{1 \perp 0}}
\def\Pperp{P_{1 \perp 0}}
\def\Pperpb{\Pperp^{\text{B}}}
\def \hatd { \widehat{ d } }
\def \dhatin {\hatd^{\mbox{\tiny IN}}}
\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
\def \mvn{\text{MVN}}
\def\mn{\text{MN}}
\def \iw{\text{W}^{-1}}
\def \fp {fastPHASE\ }
\def\var{\rm Var}
\def\mvr{\text{BMVR}}
\def\bfuni{\text{BF}_\text{uni}}
\def\one{\mbf{1}}
\def\zero{\mbf{0}}
\def\hall{H_{\text{all}}}
\def\pall{\p_{\text{all}}}
\def\bfall{\text{BF}_\text{all}}
\def\bfav{\text{BF}_\text{av}}
\def\gall{\gamma_\text{all}}
\def\limbfall{\bfall^\rightarrow}
\def\limbfav{\bfav^\rightarrow}
\def\limbfgamma{\BF_\gamma^\rightarrow}
\def\sxx{S_{xx}}
\def\syy{S_{yy}}
\def\vyy{V_{yy}}
\def\vyx{V_{yx}}
\def\vxx{{V_{xx}}}
\def\rss{\text{\rm RSS}}
\def\dss{\text{\rm $\Delta$SS}}
\def\U{U}
\def\D{D}
\def\I{I}
\def\luu{L_{UU}}
\def\indep{\perp}
\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
                                % mode too
\newcommand\height{{\it height}}
\newcommand\weight{{\it weight}}
\def\BF{\text{BF}}

@@

 The two main assumptions we will make here are that 
* these effects are exchangeable: that is (by de Finetti's theorem) they can be
thought of as being independent and identically distributed from some unknown distribution $g(\beta)$. 
* these effects are symmetrically distributed about 0, with density monotonically decaying as we move away from 0.

Specifically we will assume a parameteric form for $g$
as a mixture of 0-centered normal distributions

@@

\subsection{The False Sign Rate}

The usual definition of the local False Discovery Rate (lfdr) for observation $j$ is
$\text{lfdr}_j = p(\beta_j = 0 | \hat\beta, s)$.
The lfdr terminology comes from using "discovery" to refer to rejecting the null ($H_j:\beta_j=0$), so lfdr 
gives the probability that we make a mistake if we reject
$H_j$, that is the probability that, if we reject $H_j$,
it is a  "false discovery".

As pointed out by \cite{gelman.etal}, there are settings where
$\beta_j=0$ is implausible, in which case the lfdr is not
a useful concept: if every $\beta_j$ is non-zero then there
is no such thing as a false discovery and the lfdr will be 0. Gelman et al suggest that in such settings we
might replace the concept of a false discovery with the
concept of an "error in sign". The idea is that, in settings where $\beta_j=0$ is implausible, the most fundamental
inference objective is to ask which $\beta_j$ are positive and which are negative. Indeed, even in settings where some
$\beta_j$ are exactly zero, it could be argued that identifying which are positive and which negative is
fundamentally more interesting and useful than identifying which are non-zero. However, in such settings this usually comes down to something similar in practice: the observations that one can tell are non-zero are also the ones whose
sign can be reliably inferred. In this sense the
local false sign rate is a natural generalization of
the local false discovery rate.

Following these ideas of \cite{gelman.etal}, we define the local False Sign Rate (lfsr) as 
$$\text{lfsr}_j = \min[ p(\beta_j \geq 0| \hat\beta, s), p(\beta_j \leq 0| \hat\beta, s) ].$$
To give the intuition behind this, suppose for concreteness
that the minimum is achieved by the first term, $p(\beta_j \geq 0| \hat\beta, s)=0.05$ say. Then, if we were to declare that $\beta_j$ is negative, the probability that we have
made an error in sign would be 0.05.

Note that $\text{lfsr}_j \geq \text{lfdr}_j$ 
because both the events $\beta_j \geq 0$
and $\beta_j \leq 0$ include the event $\beta_j=0$.
Thus, lfsr gives an upper bound for the lfdr,
and so can be used as a conservative estimate of the lfdr
if an lfdr is desired. (This may be helpful when comparing
with methods that compute an lfdr.)
In addition, we intuitively expect the difference $\text{lfsr}_j - \text{lfdr}_j$ to be small, at least when $\text{lfdr}_j$ is small, because when we are confident that a particular effect is non-zero we would expect to be similarly confident about the sign of the effect.

>>>>

\def\lfdr{\text{lfdr}}
\def\lfsr{\text{lfsr}}
\def\x{\mbf{x}}
\def\y{\mbf{y}}
\def\e{\mbf{e}}
\def\g{\mbf{g}}
\def\p{p}
\def\Yb{\hat{Y}_1^{\text{BAYES}}}
\def\Xperp{X_{1 \perp 0}}
\def\Pperp{P_{1 \perp 0}}
\def\Pperpb{\Pperp^{\text{B}}}
\def \hatd { \widehat{ d } }
\def \dhatin {\hatd^{\mbox{\tiny IN}}}
\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
\def \mvn{\text{MVN}}
\def\mn{\text{MN}}
\def \iw{\text{W}^{-1}}
\def \fp {fastPHASE\ }
\def\var{\rm Var}
\def\mvr{\text{BMVR}}
\def\bfuni{\text{BF}_\text{uni}}
\def\one{\mbf{1}}
\def\zero{\mbf{0}}
\def\hall{H_{\text{all}}}
\def\pall{\p_{\text{all}}}
\def\bfall{\text{BF}_\text{all}}
\def\bfav{\text{BF}_\text{av}}
\def\gall{\gamma_\text{all}}
\def\limbfall{\bfall^\rightarrow}
\def\limbfav{\bfav^\rightarrow}
\def\limbfgamma{\BF_\gamma^\rightarrow}
\def\sxx{S_{xx}}
\def\syy{S_{yy}}
\def\vyy{V_{yy}}
\def\vyx{V_{yx}}
\def\vxx{{V_{xx}}}
\def\rss{\text{\rm RSS}}
\def\dss{\text{\rm $\Delta$SS}}
\def\U{U}
\def\D{D}
\def\I{I}
\def\luu{L_{UU}}
\def\indep{\perp}
\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
                                % mode too
\newcommand\height{{\it height}}
\newcommand\weight{{\it weight}}
\def\BF{\text{BF}}

@@

 The two main assumptions we will make here are that 
* these effects are exchangeable: that is (by de Finetti's theorem) they can be
thought of as being independent and identically distributed from some unknown distribution $g(\beta)$. 
* these effects are symmetric and unimodal about 0.

Specifically we will assume a parametric form for $g$
as a mixture of 0-centered normal distributions

@@

\subsection{The local False Sign Rate}

The local False Discovery Rate (lfdr) for observation $j$ is
\begin{equation}
\lfdr_j = p(\beta_j = 0 | \hat\beta, s).
\end{equation}
The lfdr terminology comes from using "discovery" to refer to rejecting the null ($H_j:\beta_j=0$).
Specifically $\lfdr_j$ is the probability that, if we reject $H_j$,
it is a  "false discovery".

As pointed out by \cite{gelman.etal}, there are settings where
$\beta_j=0$ is implausible, in which case the lfdr is not
a useful concept: if every $\beta_j$ is non-zero then there
is no such thing as a false discovery and the lfdr will be identically 0 for all methods. 
Gelman et al suggest that in such settings we
might replace the concept of a false discovery with the
concept of an "error in sign". The idea is that, in settings where $\beta_j=0$ is implausible, the most fundamental
inference objective is to ask which $\beta_j$ are positive and which are negative. Indeed, even in settings where some
$\beta_j$ are exactly zero, it could be argued that identifying which are positive and which negative is
fundamentally more interesting and useful than identifying which are non-zero.

Following these ideas of \cite{gelman.etal}, we define the local False Sign Rate (lfsr) for observation $j$ as 
\begin{equation}
\lfsr_j := \min[ p(\beta_j \geq 0| \hat\beta, s), p(\beta_j \leq 0| \hat\beta, s) ].
\end{equation}
Thus $\lfsr_j$ gives the probability that we would get the sign of $\beta_j$ wrong if we
were to make our best guess. (Note that we count it as an error to state that $\beta$ is positive or negative when it is truly zero.)
To illustrate, suppose for concreteness
that the minimum is achieved by the first term, $p(\beta_j \geq 0| \hat\beta, s)=0.05$ say. Then
our best guess would be that $\beta$ is negative, and the probability that we have
made an error in sign would be 0.05.

Note that $\text{lfsr}_j \geq \text{lfdr}_j$ 
because both the events $\beta_j \geq 0$
and $\beta_j \leq 0$ include the event $\beta_j=0$.
Thus, lfsr gives an upper bound for lfdr.

### 2013-11-07 talks/NSmeet2013.Rmd
add some R code for paper, and update some of the options

# Example: BRCA1 vs BRCA2 expression

@@

```{r, echo=FALSE,fig.height=3.5,fig.width=5,fig.cap=""}
  nc=40 #number of bins in histogram
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

@@

# Example: BRCA1 vs BRCA2 expression

```{r, echo=FALSE,fig.height=3.5,fig.width=5,fig.cap=""}
  nc=40 #number of bins in histogram
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

@@

# Example: BRCA1 vs BRCA2 expression

```{r, echo=FALSE,fig.height=3.5,fig.width=5,fig.cap=""}
  nc=40 #number of bins in histogram
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

@@

# Example 2: Mouse Heart Data

@@

# Example 2: Mouse Heart Data

@@

# FDR problem 1: different genes have different precision/power

```{r, echo=FALSE,fig.height=4,fig.cap=""}
hist(log10(xx),main="Counts vary considerably across genes", xlab="log10(counts)")
```

# FDR problem 1: lower count genes, less power, add noise

@@

# FDR problem 1: higher count genes, more power
```{r, echo=FALSE, fig.height=4,fig.cap=""}

@@

abline(h=qval.high$pi0,col=nullcol,lwd=2)
h$density=rep(qval.high$pi0,length(h$density))  
plot(h,col=nullcol,add=TRUE,freq=FALSE)
abline(v=0.05,lwd=3)
text(0.6,3,labels=paste0("FDR = ",round(max(qval.high$qval[pval.high<0.05]),2)),cex=1.2)
#qval.high$q[pval.high<0.05] gives FDR for p<0.05
```

# FDR problem 1: q values increased by low count genes
```{r,echo=FALSE, fig.height=4,fig.cap=""}
plot(qval$qval[xx>1000],qval.high$qval,xlab="q values from all gene analysis", ylab="q values from high-count gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
abline(a=0,b=1,col=2)
```

@@

- This allows us to estimate $\pi_0$ ``conservatively" using the density of $p$ values near 1.

@@

# Proposed Alternative

- Instead of working with $z$ scores or $p$ values, work with two numbers 
$(\hat\beta_j,s_j)$ for each test.

- *Incorporate the precision* of the observations $\hat\beta$ into the likelihood.
Specifically, we approximate likelihood for $\beta_j$ by a normal (``Laplace") approximation: 
$$L(\beta_j) \propto \exp(-0.5 (\beta_j - \hat\beta_j)^2/s_j^2).$$
[From $\hat\beta_j \sim N(\beta_j, s_j)$]

- *Directly model the underlying distribution of $\beta$*, using a unimodal family of distributions $g$; estimate $g$ from the data.

# Proposed Alternative: More details

@@

- The estimating $g$ comes down to estimating $\pi$ - easily done
by maximum likelihood (EM algorithm) or variational Bayes.

@@

- Alternatively, a mixture of uniforms, with 0 as one end-point of the range,
provides still more flexibility, and in particular allows for asymmetry. (Grenander 1953 shows this provides the non-parametric mle in the ``no error" case; Campy + Thomas do the equal-errors case)

@@

```{r}
hh.ash = ash(hh.betahat,hh.sebetahat)
hh.ash.hu = ash(hh.betahat, hh.sebetahat,mixcompdist="halfuniform")
```

```{r, echo=FALSE,fig.height=4,fig.cap=""}
h=hist(hh.betahat,freq=FALSE,main="BRCA, effect sizes",ylim=c(0,2),nclass=nc)
x = seq(-3,3,length=100)
lines(x,density(hh.ash,x),col=2,lwd=2)
lines(x,density(hh.ash.hu,x),col=3,lwd=2)
```

# Illustration: BRCA data
```{r, echo=FALSE,fig.height=4,fig.cap=""}
nullalthist(hh.zscore,hh.ashz$ZeroProb,main="ash")
```

# Illustration: BRCA data
```{r, echo=FALSE,fig.height=4,fig.cap=""}
nullalthist(hh.zscore,hh.mixfdr$fdr,main="mixfdr")
```

@@

- Similarly, the data can't tell us what proportion of g should be exactly on zero, vs near zero

- Of course this is true regardless of the method used!

@@

@@

# Estimation

- This approach also provides a full posterior distribution for each $\beta_j$.

- So for example we can easily compute fdrs for discoveries other than ``non-zero" (eg genes with at least 2-fold difference between conditions

Compare fit of $g$ both allowing $\beta_j=0$ and not allowing it.
```{r, echo=FALSE, includ=FALSE}
  hh.ash.fdr = ash(hh.betahat,hh.sebetahat,usePointMass = TRUE,prior="nullbiased",VB=TRUE)
  hh.ash.shrink = ash(hh.betahat,hh.sebetahat)
```

@@

- Because $\pi$ is estimated from the data, the amount
of shrinkage is adaptive to the data. And because of the role of $s_j$, the amount of shrinkage adapts to the information on each gene.

@@

```{r, echo=FALSE,fig.height=3,fig.width=3,fig.cap=""}
plot(zdat.ash$localfdr,pval,ylab="p value", xlab="ASH local fdr")
```

@@

```{r, echo=FALSE,fig.height=3,fig.width=3,fig.cap=""}
plot(zdat.ash$localfdr,pval,ylab="p value", xlab="ASH local fdr")
points(zdat.ash$localfdr[15325],pval[15325],col=2,pch=16)
points(zdat.ash$localfdr[16123],pval[16123],col=2,pch=16)
```

@@

```{r, echo=FALSE}
cbind(x[,1:5],pval,zdat.ash$localfdr)[c(15325,16123),]
```

# Compare q values for high count genes, with and without low count genes

```{r, echo=FALSE,fig.height=3,fig.width=3,fig.cap=""}
plot(zdat.ash$qvalue[xx>1000],zdat.ash.high$qvalue,xlab="q values from all data",ylab="q values from high only")
abline(a=0,b=1,col=2,lwd=2)
```

# BRCA1: Compare $\pi_0$ estimates
```{r}
  c(hh.fdrtool$param[3],
  hh.locfdr$fp0[1,3],
  hh.mixfdr$pi[1],
  hh.ashz$fitted.g$pi[1])
```

@@

```{r, echo=FALSE,fig.height=3,fig.width=3}
  plot(zdat[3,],zdat[4,],ylab="standard error",xlab="beta-hat")
  x[tail(order(zdat[3,])),]
```

>>>>

# FDR, local fdr, and q values

Although precise definitions vary depending on whether one
takes a Bayesian or Frequentist approach to the problem, roughly

- The FDR at a threshold $P$ is 
$$\text{FDR}(P)=\Pr(\beta_j = 0 |  p_j<P).$$

- The q value for observation $j$ is $q_j=\text{FDR}(p_j)$.

- The local false discovery rate, fdr, at threshold $P$ is 
$$\text{fdr}(P) = \Pr(\beta_j =0 | p_j=P ).$$

- The fdr is more relevant, but slightly harder to estimate than 
FDR because it involves density estimation rather than tail-area estimation.

# Example: FDR estimation

@@

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  nc=40 #number of bins in histogram
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

@@

# Example:  FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  nc=40 #number of bins in histogram
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

@@

  hh.hist$density=rep(hh.q$pi0,length(hh.hist$density))  
  #hh.hist$counts=rep(hh.q$pi0*length(hh.pval)/nc,length(hh.hist$counts)) 
  plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)
```

Data from Hedenfalk et al, comparing BRCA1 vs BRCA2 expression.

# Example:  FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  nc=40 #number of bins in histogram
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

@@

# Example: fdr estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  nc=40 #number of bins in histogram
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

# Example: fdr estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  nc=40 #number of bins in histogram
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

  require(fdrtool)
  hh.gren = grenander(ecdf(hh.pval))
  abline(h=min(hh.gren$f.knots),col=nullcol,lwd=2)  
  lines(hh.gren$x.knots,hh.gren$f.knots,lwd=2)
```

# Example: fdr estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  nc=40 #number of bins in histogram
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

  require(fdrtool)
  hh.gren = grenander(ecdf(hh.pval))
  abline(h=min(hh.gren$f.knots),col=nullcol,lwd=2)  
  lines(hh.gren$x.knots,hh.gren$f.knots,lwd=2)
  abline(v=0.1,lwd=2,col=2)
  text(0.1,0.9,labels="a",col=2,cex=1)  
  text(0.1,0.34,labels="b",col=2,cex=1.2)  
  text(0.6,3,labels=paste0("fdr = b/(a+b) =  ",round(min(hh.gren$f.knots)/approx(hh.gren$x.knots,hh.gren$f.knots,0.1)$y,2)),cex=1.2)
```

# FDR problem 1: different measurement precision

- If some effects are measured very imprecisely, those tests ``lack power"
and simply add noise

- In particular, such tests increase the estimated number of nulls, and increase
the FDR for other tests

- It would seem preferable to simply ignore the tests with very low precision. Summarizing each test by a $p$ value (or $Z$ score) loses the information about precision.

# Example: Mouse Heart Data

@@

# Example: Mouse Heart Data

@@

# Mouse Data: Counts vary considerably across genes

```{r, echo=FALSE,fig.height=4,fig.cap=""}
hist(log10(xx),main="Distribution of total counts", xlab="log10(counts)")
```

# Lower count genes, less power

@@

# Higher count genes, more power
```{r, echo=FALSE, fig.height=4,fig.cap=""}

@@

abline(h=qval.high$pi0,col=nullcol,lwd=2)
h$density=rep(qval.high$pi0,length(h$density))  
plot(h,col=nullcol,add=TRUE,freq=FALSE)
abline(v=0.05,lwd=3)
#text(0.6,3,labels=paste0("FDR = ",round(max(qval.high$qval[pval.high<0.05]),2)),cex=1.2)
#qval.high$q[pval.high<0.05] gives FDR for p<0.05
```

# FDR problem 1: low count genes add noise, increase q values
```{r,echo=FALSE, fig.height=4,fig.cap=""}
plot(qval.high$qval,qval$qval[xx>1000],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
abline(a=0,b=1,col=2)
```

# FDR problem 1: Summary

- Analyzing $p$ values or $Z$ scores doesn't fully account 
for measurement precision.

@@

- The ZA allows us to estimate the null proportion, $\pi_0$, using the density of $p$ values near 1 (or $Z$ scores near 0).

@@

# FDR via Empirical Bayes

- Following previous work (e.g. Newton, Efron, Muralidharan) we take an empirical Bayes approach to FDR.

- Eg Efron assumes that the $Z$ scores come from a mixture of null, and alternative:
$$Z_j \sim f_Z(.) = \pi_0 N(.;0,1) + (1-\pi_0) f_1(.)$$
where $f_1$ is to be estimated from the data.

- Various semi-parametric approaches taken to estimating $f_1$. For example,
Efron uses Poisson regression; Muralidharan uses mixture of normal distributions.

- $\text{fdr}(Z) \approx \pi_0 N(Z; 0,1)/ f_Z(Z)$

# FDR: The New Deal

- Instead of modelling $Z$ scores, model the effects $\beta$,
$$\beta_j \sim \pi_0 \delta_0(.) + (1-\pi_0) g(.)$$

- Constrain $g$ to be unimodal about 0.

- *Incorporate precision* of each observation $\hat\beta$ into the likelihood.
Specifically, approximate likelihood for $\beta_j$ by a normal: 
$$L(\beta_j) \propto \exp(-0.5 (\beta_j - \hat\beta_j)^2/s_j^2).$$
[From $\hat\beta_j \sim N(\beta_j, s_j)$]

- fdr given by $$p(\beta_j =0 | \hat\beta_j) = \pi_0 p(\hat\beta_j | \beta_j=0)/p(\hat\beta_j)$$

# FDR - The New Deal

@@

- Estimating $g$ comes down to estimating $\pi$. Joint estimation of $\pi_0,\pi$ easy by maximum likelihood (EM algorithm) or variational Bayes.

@@

- Alternatively, a mixture of uniforms, with 0 as one end-point of the range,
provides still more flexibility, and in particular allows for asymmetry.

- If allow a very large number of uniforms this provides the non-parametric mle for $g$; cf Grenander 1953; Campy + Thomas.

@@

@@

- As a result $\pi_0$ is formally unidentifiable. Eg data can never rule out $\pi_0=0$.

@@

```{r, echo=FALSE,include=FALSE}
hh.ash = ash(hh.betahat,hh.sebetahat)
hh.ashz = ash(hh.zscore,1, pointmass=TRUE, prior="nullbiased")
hh.ash.hu = ash(hh.betahat, hh.sebetahat,mixcompdist="halfuniform")
```

Compare fitted $f(\beta)$, both estimating $\pi_0$ and fixing $\pi_0=0$.
```{r, echo=FALSE, includ=FALSE}
  hh.ash.fdr = ash(hh.zscore,1,pointmass = TRUE,prior="nullbiased")
  hh.ash.shrink = ash(hh.zscore,1)
```

# Recall Problem: distribution of alternative Z values multimodal
```{r, echo=FALSE,fig.height=4,fig.cap=""}
nullalthist(hh.zscore,hh.mixfdr$fdr,main="mixfdr")
```

# Problem Fixed: distribution of alternative Z values unimodal
```{r, echo=FALSE,fig.height=4,fig.cap=""}
nullalthist(hh.zscore,hh.ashz$ZeroProb,main="ash")
```

# BRCA1: Compare $\pi_0$ estimates
```{r}
  round(c(hh.fdrtool$param[3],
  hh.locfdr$fp0[1,3],
  hh.mixfdr$pi[1],
  hh.ashz$fitted.g$pi[1]),2)
```

@@

@@

# Estimation and Shrinkage

- Besides allowing one to estimate fdr and fsr, 
this approach also provides a full posterior distribution for each $\beta_j$.

- So for example we can easily compute fdrs for discoveries other than ``non-zero" (eg compute $\Pr(|\beta_j| > 2 \| \hat\beta_j)$).

- And use it to obtain point estimates and credible intervals for each $\beta_j$, taking account of information from all the other $\beta_j$.

- Because $f(\beta)$ is unimodal, the point estimates will tend to be ``shrunk" towards the overall mean (0).

- Because $f(\beta)$ is estimated from the data, the amount
of shrinkage is adaptive to the data. And because of the role of $s_j$, the amount of shrinkage adapts to the information on each gene.

- So we call the approach ``Adaptive Shrinkage" (ASH).

@@

```{r, echo=FALSE,fig.height=3,fig.cap=""}
plot(zdat.ash$localfdr,pval,ylab="p value", xlab="ASH local fdr")
```

@@

```{r, echo=FALSE,fig.height=3,fig.cap=""}
plot(zdat.ash$localfdr,pval,ylab="p value", xlab="ASH local fdr")
points(zdat.ash$localfdr[15325],pval[15325],col=2,pch=16)
points(zdat.ash$localfdr[16123],pval[16123],col=2,pch=16)
```

@@

```{r, echo=FALSE}
x = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
xx = rowSums(x[,2:5])
x = x[xx>0,]
xx = xx[xx>0]
cbind(x[,1:5],pval,zdat.ash$localfdr)[c(15325,16123),]
```

# Recall FDR problem 1: q values increased by low count genes
```{r,echo=FALSE, fig.height=4,fig.cap=""}
plot(qval.high$qval,qval$qval[xx>1000],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
abline(a=0,b=1,col=2)
```

# ASH q values more robust to inclusion of low count genes

```{r, echo=FALSE,fig.height=3,fig.cap=""}
plot(zdat.ash.high$qvalue,zdat.ash$qvalue[xx>1000],xlab="q values from high only",ylab="q values from all data")
abline(a=0,b=1,col=2,lwd=2)
```

@@

```{r, echo=FALSE,fig.height=3}
  plot(zdat[3,],zdat[4,],ylab="standard error",xlab="beta-hat")
  x[tail(order(zdat[3,])),]
```
